### 2020-03-09 parenting_obs_e2/joint_attention/parenting_obs_JA.Rmd

---
title: "Parenting Observational Study E2 Coding"
author: "Emily, Vivian, George"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    number_sections: true
---

<style type="text/css">
body, td {
   font-size: 14px;
}
code {
  font-size: 11px;
}
pre {
  font-size: 11px;
}
</style>

Looking at the effect of electronically-delivered parenting interventions on parent-child interactions. Parents in the experimental condition viewed a video (designed for a parenting app) demonstrating an activity they could do with their child. Parents in the control condition did not view a video. Both groups of parents were asked to play with their child for 3 minutes with a set of props that matched the video demonstration they had seen. Children were 6-24 months of age. There were 6 videos total (2 per age group). Assignment to videos/sets of toys was counterbalanced within age groups. Parents also completed the Parenting Attitudes Questionnaire. Videos of the play sessions were coded for bids for joint attention, episodes of passive joint attention, and episodes of coordinated joint attention.

# Data preprocessing

Preliminaries.

```{r echo=FALSE}
rm(list=ls())
knitr::opts_chunk$set(fig.width=8, fig.height=5, 
                      echo=TRUE, warning=FALSE, message=FALSE, cache=TRUE)
library(tidyverse)
library(lme4)
library(knitr)
library(markdown)
library(lmerTest)
library(stringr)
library(langcog)
library(xtable)
library(Hmisc)
theme_set(ggthemes::theme_few() + theme(legend.title = element_text(size=18), legend.text = element_text(size=16), 
        axis.text.x  = element_text(vjust=0.5, size=16), axis.text.y  = element_text(vjust=0.5, size=16),
        axis.title.x = element_text(size=18), axis.title.y = element_text(size=18)))
```

Read in files and consolidate to the same directory.

```{r echo=F}
combine_datavyu_files <- function(data_dir) {
  d.long <- data.frame()
  files <- dir(data_dir)

  for (f in files) {
    jf <- paste(data_dir, f, sep="")
    jd <- read.csv(jf)

  # (assumes something happened early in the vid and late in the vid..)
    vid_duration = jd[nrow(jd),]$JointAttention.offset - jd[1,]$JointAttention.onset
    if(vid_duration < 180000) print(paste(f, "final - first coded time is less than 180 s:",vid_duration))

    jd$phase_length = jd$JointAttention.offset - jd$JointAttention.onset
    if(length(which(jd$phase_length < 0))!=0) print(paste(f, "episode with negative duration!"))
    which(jd$phase_length < 3000) # trim episodes <3 seconds long?

    id <- data.frame(SID = str_replace(f, ".csv", ""), # "012017_01_recode"
                   line = jd$JointAttention.ordinal, 
                   phase_length = jd$phase_length,
                   pja = jd$JointAttention.pJA, # was episode passive?
                   cja = jd$JointAttention.cJA,
                   bid = jd$JointAttention.bids, # parent bids (1=parent, 0=kid)
                   vid_duration = vid_duration
                )

    d.long <- rbind(d.long, id)
    #d.raw <- bind_rows(d.raw, sd)
  }
  return(d.long)
}

d.long <- combine_datavyu_files("data/")

d.long[which(d.long$phase_length<3000),c("cja","pja")] = 0 # any episodes <3s not counted (except bids)

# summarise per subject
d.raw <- d.long %>% group_by(SID) %>%
  mutate(vid_duration = vid_duration / 60000, # seconds to minutes
         phase_length = phase_length / 60000) %>%
  summarise(bids=sum(bid), pja_length=sum(pja*phase_length), cja_length=sum(cja*phase_length),
            pja=sum(pja), cja=sum(cja), vid_duration=vid_duration[1]) %>%
  mutate(total_ja = pja + cja, total_lja = pja_length + cja_length)

# normalize to rates (e.g., bids/minute)
d.raw[,2:9] = d.raw[,2:9] / d.raw$vid_duration
```

Read in trial info 
```{r}
conditions <- read.csv("conditions.csv") %>%
  transmute(SID = sid, condition=condition, video=video, label=label)

d <- left_join(d.raw, conditions, by="SID")

#d$age[d$sid=="111517_3"] = "2"
```

# Reliability

Import reliability coder's data (20 participants were coded for reliability).
```{r}
d.long_rel <- combine_datavyu_files("rel_data/")

d.long_rel[which(d.long_rel$phase_length<3000),c("cja","pja")] = 0 # any episodes <3s not counted (except bids)

# summarise per subject
d.raw_rel <- d.long_rel %>% group_by(SID) %>%
  mutate(vid_duration = vid_duration / 60000, # seconds to minutes
         phase_length = phase_length / 60000) %>%
  summarise(bids=sum(bid), pja_length=sum(pja*phase_length), cja_length=sum(cja*phase_length),
            pja=sum(pja), cja=sum(cja), vid_duration=vid_duration[1]) %>%
  mutate(total_ja = pja + cja, total_lja = pja_length + cja_length)

# normalize to rates (e.g., bids/minute)
d.raw_rel[,2:9] = d.raw_rel[,2:9] / d.raw_rel$vid_duration

```

Reliability.
```{r}
d.raw_rel <- d.raw_rel %>%
  mutate(bids_rel = bids, pja_rel = pja, cja_rel = cja, l_pja_rel = pja_length, l_cja_rel = cja_length) %>%
  select(SID, bids_rel, pja_rel, cja_rel, l_cja_rel, l_pja_rel)

d.raw <- d.raw %>% mutate(l_pja = pja_length, l_cja = cja_length)
# 110416_03 l_cja_rel = -25889, which shouldn't be possible (incomplete final row of file)
# 121516_01 has NA bids_rel, pja_rel, and cja_rel..
# 111716_08 l_cja = -49923, which shouldn't be possible

# 110416_05, 050116_01, and 092216_03: reliability coder saw no passive JA
d_rel <- d.raw_rel %>% left_join(d.raw, by="SID") %>%
  select(SID, bids, bids_rel, pja, pja_rel, cja, cja_rel, l_cja, l_cja_rel, l_pja, l_pja_rel)

get_ICC2 <- function(dat)  {
  rel = psych::ICC(dat, missing = FALSE, alpha = 0.05)
  return(rel$results[2,])
}

bids_rel = get_ICC2(d_rel[,c("bids","bids_rel")]) # .85

pja_rel = get_ICC2(d_rel[,c("pja","pja_rel")]) # .75 eps pass JA

cja_rel = get_ICC2(d_rel[,c("cja","cja_rel")]) # .75 eps coord JA

l_pja_rel = get_ICC2(d_rel[,c("l_pja","l_pja_rel")]) # .73 pass JA length

l_cja_rel = get_ICC2(d_rel[,c("l_cja","l_cja_rel")]) # .75 coord JA length
```

Inter-rater reliability is very high for number of bids, high for number of episodes of coordinated joint attention, but there is less agreement for episodes passive joint attention.

# Plots

Total number of bids (successful and unsuccessful)
```{r, width=4, height=4}
dem <- read.csv("../anonymized_demographics_parenting_obs_e2.csv", header=T)
dem$age = as.numeric(as.character(dem$age))
d$age = NULL
d <- d %>% left_join(dem, by="SID") %>%
  filter(age <= 2.0) # 1 kid exactly 2...

save(d, file="Exp2_joint_attention_data.RData")

d$condition <- factor(d$condition, levels = c("con", "exp"), labels = c("control", "video"))

ms_bids <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "bids")

ggplot(ms_bids, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Total bids for Joint Attention") +
  langcog::scale_colour_solarized() + theme(legend.position="none") 
```

Total episodes of joint attention.

```{r}
ms_tja <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "total_ja")

ggplot(ms_tja, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total episodes of Joint Attention") +
  langcog::scale_colour_solarized() 
```

Episodes of passive joint attention.
```{r}
ms_pja <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "pja")

ggplot(ms_pja, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Episodes of Passive Joint Attention")  +
  langcog::scale_colour_solarized()
```

Episodes of coordinated joint attention
```{r}
ms_cja <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "cja")

ggplot(ms_cja, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Episodes of Coordinated Joint Attention")  +
  langcog::scale_colour_solarized()
```

Total duration of passive joint attention.
```{r}
ms_lpja <- d %>%
  filter(!is.na(condition))%>%
  group_by(condition) %>%
  multi_boot_standard(col = "pja_length")

ggplot(ms_lpja, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total Duration Passive Joint Attention")  +
  langcog::scale_colour_solarized() 
```

Total duration of coordinated joint attention.
```{r}
ms_lcja <- d %>%
  filter(!is.na(condition))%>%
  group_by(condition) %>%
  multi_boot_standard(col = "cja_length")

ggplot(ms_lcja, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total Duration Coordinated Joint Attention")  +
  langcog::scale_colour_solarized() 
```

##Demographics

###Child gender

Total bids by gender.
```{r}
# total bids
ms_bids <- d %>%
  group_by(condition, gender) %>%
  multi_boot_standard(col = "bids")

ggplot(ms_bids, aes(x = condition, y = mean, fill=gender)) + 
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total bids for Joint Attention")  +
  langcog::scale_colour_solarized() 
```

Total duration of passive joint attention by gender
```{r}
ms_lpja <- d %>%
  filter(!is.na(condition))%>%
  group_by(condition, gender) %>%
  multi_boot_standard(col = "pja_length")

ggplot(ms_lpja, aes(x = condition, y = mean, fill=gender)) + 
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total Duration Passive Joint Attention")  +
  langcog::scale_colour_solarized() 
```

Total duration of coordinated joint attention by gender
```{r}
ms_lcja <- d %>%
  filter(!is.na(condition))%>%
  group_by(condition, gender) %>%
  multi_boot_standard(col = "cja_length")

ggplot(ms_lcja, aes(x = condition, y = mean, fill=gender)) + 
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total Duration Coordinated Joint Attention")  +
  langcog::scale_colour_solarized() 
```

###Child age

Children were separated into 3 age groups: 6-12 months, 12-18 months, and 18-24 months.

Total bids by age.
```{r}
# total bids
d$age_group = ifelse(d$age<1, "6-12", ifelse(d$age<1.5, "12-18", "18-24"))
d$age_group <- factor(d$age_group , levels = c("6-12","12-18", "18-24"))

ms_bids <- d %>%
  group_by(condition, age_group) %>%
  multi_boot_standard(col = "bids")

ggplot(ms_bids, aes(x = condition, y = mean, fill=age_group)) + 
  geom_bar(stat="identity", position= "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total bids for Joint Attention")  +
  langcog::scale_colour_solarized() 
```

Total duration of passive joint attention by age
```{r}
ms_lpja <- d %>%
  filter(!is.na(condition))%>%
  group_by(condition, age_group) %>%
  multi_boot_standard(col = "pja_length")

ggplot(ms_lpja, aes(x = condition, y = mean, fill = age_group)) + 
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total Duration Passive Joint Attention")  +
  langcog::scale_colour_solarized() 
```

Total duration of coordinated joint attention by age
```{r}
ms_lcja <- d %>%
  filter(!is.na(condition))%>%
  group_by(condition, age_group) %>%
  multi_boot_standard(col = "cja_length")

ggplot(ms_lcja, aes(x = condition, y = mean, fill = age_group)) + 
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Total Duration Coordinated Joint Attention")  +
  langcog::scale_colour_solarized() 
```

# Analysis

Is the duration of joint attention correlated with age?

```{r}
cormat <- d %>%
  select(age, pja_length, cja_length, total_lja)
rcorr(as.matrix(cormat))
```

Mixed effects models.

Prepare data.
```{r}
lmer_data <- d %>% 
  mutate(Condition = factor(condition), 
         bids = as.numeric(bids),
         age = as.numeric(langcog::scale(age, scale=FALSE)),
         gender = as.factor(gender),
         Video = as.factor(video))
```

Total number of bids
```{r}
lmer_data$parent_ed = as.numeric(as.character(lmer_data$parent_ed))
maximal_mod <- lmer(bids ~ Condition * age + (1| Video), data = lmer_data)
summary(maximal_mod)
```

Episodes of coordinated joint attention.
```{r}
maximal_mod <- lmer(cja ~ Condition * age + (1| Video), data = lmer_data)
summary(maximal_mod)
```

Episodes of passive joint attention.
```{r}
maximal_mod <- lmer(pja ~ Condition * age + (1| Video), data = lmer_data)
summary(maximal_mod)
```

Total duration of passive joint attention.
```{r}
maximal_mod <- lmer(pja_length ~ Condition * age + (1| Video), data = lmer_data)
summary(maximal_mod)
```

Total duration of coordinated joint attention.
```{r}
maximal_mod <- lmer(cja_length ~ Condition * age + (1| Video), data = lmer_data)
summary(maximal_mod)
```

# Conclusions