### 2020-03-16 parenting_obs_e1/joint_attention/parenting_obs_JA.Rmd
switched to rate of JA per min

---
title: "Parenting Observational Study E1"
author: "Emily & George"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    number_sections: true
---

<style type="text/css">
body, td {
   font-size: 14px;
}
code {
  font-size: 11px;
}
pre {
  font-size: 11px;
}
</style>

Looking at the effect of electronically-delivered parenting interventions on parent-child interactions. Parents in the experimental condition viewed a video (designed for a parenting app) demonstrating an activity they could do with their child. Parents in the control condition did not view a video. Both groups of parents were asked to play with their child for 3 minutes with a set of props that matched the video demonstration they had seen. Children were 6-24 months of age. There were 6 videos total (2 per age group). Assignment to videos/sets of toys was counterbalanced within age groups. Parents also completed the Parenting Attitudes Questionnaire. Videos of the play sessions were coded for bids for joint attention, episodes of passive joint attention, and episodes of coordinated joint attention.

# Data preprocessing

Preliminaries.

```{r include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(fig.width=8, fig.height=5, 
                      echo=TRUE, warning=FALSE, message=FALSE, cache=TRUE)
library(tidyverse)
library(lme4)
library(knitr)
library(markdown)
library(lmerTest)
library(stringr)
library(langcog)
library(xtable)
library(Hmisc)
theme_set(ggthemes::theme_few() + theme(legend.title = element_text(size=18), legend.text = element_text(size=16), 
        axis.text.x  = element_text(vjust=0.5, size=16), axis.text.y  = element_text(vjust=0.5, size=16),
        axis.title.x = element_text(size=18), axis.title.y = element_text(size=18)))
```

Read in files and consolidate to the same directory.

```{r}
# for rate coding (JA/minute), we coded the start / end of play sessions because they were not exactly 3 minutes
play_duration <- read.csv("parenting_obs_play_duration_E1.csv", header=T)
play_duration <- play_duration %>% 
  mutate(play_duration = play_stop_sec - play_start_sec)
sort(play_duration$play_duration) # shortest: 167 seconds
summary(play_duration$play_duration) # median = 201.5, mean=205.9
# use play_duration$verbalcue_start_sec instead of play_start? (when it exists)

combine_datavyu_files <- function(data_dir) {
  d.long <- data.frame()
  files <- dir(data_dir)

  for (f in files) {
    jf <- paste(data_dir, f, sep="")
    jd <- read.csv(jf)

    # (assumes something happened early in the vid and late in the vid..)
    #vid_duration = jd[nrow(jd),]$JointAttention.offset - jd[1,]$JointAttention.onset
    #if(vid_duration < 180000) print(paste(f, "final - first coded time is less than 180 s:",vid_duration))
    SID = str_replace(f, ".csv", "")
    vid_duration = play_duration[which(play_duration$SID==SID),]$play_duration

    jd$phase_length = jd$JointAttention.offset - jd$JointAttention.onset
    if(length(which(jd$phase_length < 0))!=0) print(paste(f, "episode with negative duration!"))

    id <- data.frame(SID = SID, # "012017_01_recode"
                   line = jd$JointAttention.ordinal, # numbered list of episodes e.g. 0 1 2 3 4 5 6 7
                   phase_length = jd$phase_length,
                   pja = jd$JointAttention.pJA, # was episode passive?
                   cja = jd$JointAttention.cJA,
                   bid = jd$JointAttention.bids, # parent bids (1=parent, 0=kid)
                   vid_duration = vid_duration
                )

    d.long <- rbind(d.long, id)
    #d.raw <- bind_rows(d.raw, sd)
  }
  return(d.long)
}

d.long <- combine_datavyu_files("data/")

d.long[which(d.long$phase_length<0),] # 3, but only bids so duration doesn't matter

dim(d.long %>% filter((cja==1 | pja==1) & phase_length<3000)) # 150

d.long[which(d.long$phase_length<3000),c("cja","pja")] = 0 # any episodes <3s not counted (except bids)

d.long[which(rowSums(d.long[,c("pja","cja","bid")])==0),] # now several rows are not even classified as bids - should elim

# summarise per subject
# Emily only coded unsuccessful parent bids in Exp 1, assuming that any cJA episode started with a parent bid
d.raw <- d.long %>% group_by(SID) %>%
  mutate(vid_duration = vid_duration / 60, # seconds to minutes
         phase_length = phase_length / 1000) %>% # ms to seconds
  summarise(bids=sum(bid) + sum(cja), pja_length=sum(pja*phase_length), cja_length=sum(cja*phase_length),
            pja=sum(pja), cja=sum(cja), vid_duration=vid_duration[1]) %>%
  mutate(total_ja = pja + cja, total_lja = pja_length + cja_length)

# normalize to rates (e.g., bids/minute and seconds / min duration of JA)
vid_dur = d.raw$vid_duration
d.raw[,2:9] = d.raw[,2:9] / d.raw$vid_duration
d.raw$vid_duration = vid_dur

```

Read in trial info and demographics.

```{r}
conditions <- read.csv("conditions.csv")

load("../paq/paq_demo.RData")
load("../paq/obs_demo.RData")

ids <- ids%>%
  mutate(SID = sid)%>%
  select(-sid)

demo <- demo%>%
  mutate(SID = sid)%>%
  select(-sid)%>%
  mutate(age_months = age * 12)

demo$age_group[demo$age_months < 12] <- "6"
demo$age_group[demo$age_months >= 12 & demo$age_months < 18 ] <- "12"
demo$age_group[demo$age_months >= 18] <- "18"

#note from Emily- I calculated total bids by adding unsuccessful bids + episodes of joint attention that were acheived. In retrospect, this assumes that all episodes of JA were initiated by parent, which is probably false. Next time, code all parent bids rather than only unsuccessful ones.
d <- d.raw %>%
  left_join(conditions)%>%
  mutate(bids_tot = bids + cja + pja)%>%
  left_join(demo)%>%
  left_join(ids)%>%
  select(-sid)
```

# Demographics

```{r}
demos <- conditions%>%
  left_join(demo)

demos$parent_ed[demos$parent_ed == 1] <- 8
demos$parent_ed[demos$parent_ed == 2] <- 12
demos$parent_ed[demos$parent_ed == 3] <- 14
demos$parent_ed[demos$parent_ed == 4] <- 18
demos$parent_ed[demos$parent_ed == 5] <- 20
demos$parent_ed[demos$parent_ed == 6] <- 22

#child gender
table(demos$Condition, demos$gender)
#ethnicity
table(demos$Condition, demos$ethnicity)
#years of education
table(demos$Condition, demos$parent_ed)
```

# Reliability

Import reliability coder's data (20 participants were coded for reliability).
```{r}
d.long_rel <- combine_datavyu_files("rel_data/")

d.long_rel[which(d.long_rel$phase_length<0),] 
# SID=120916_03	 line=7	 -1275	 0	1	 1	 - cja + bid, so should double-check

dim(d.long_rel %>% filter((cja==1 | pja==1) & phase_length<3000)) # 39

d.long_rel[which(d.long_rel$phase_length<3000),c("cja","pja")] = 0 # any episodes <3s not counted (except bids)

# summarise per subject
d.raw_rel <- d.long_rel %>% group_by(SID) %>%
  mutate(vid_duration = vid_duration / 60, # seconds to minutes
         phase_length = phase_length / 1000) %>% # ms to seconds
  summarise(bids=sum(bid) + sum(cja), pja_length=sum(pja*phase_length), cja_length=sum(cja*phase_length),
            pja=sum(pja), cja=sum(cja), vid_duration=vid_duration[1]) %>%
  mutate(total_ja = pja + cja, total_lja = pja_length + cja_length)

# normalize to rates (e.g., bids/minute)
d.raw_rel[,2:9] = d.raw_rel[,2:9] / d.raw_rel$vid_duration

```

## Check Reliability
```{r}
d.raw_rel <- d.raw_rel %>%
  mutate(bids_rel = bids, pja_rel = pja, cja_rel = cja, l_pja_rel = pja_length, l_cja_rel = cja_length) %>%
  select(SID, bids_rel, pja_rel, cja_rel, l_cja_rel, l_pja_rel)

d.raw %>% summarise(bids = mean(bids), cja=mean(cja), pja=mean(pja), l_pja=mean(pja_length), l_cja=mean(cja_length))

d.raw <- d.raw %>% mutate(l_pja = pja_length, l_cja = cja_length)

d_rel <- d.raw_rel %>% left_join(d.raw, by="SID") %>%
  select(SID, bids, bids_rel, pja, pja_rel, cja, cja_rel, l_cja, l_cja_rel, l_pja, l_pja_rel)

get_ICC2 <- function(dat)  {
  rel = psych::ICC(dat, missing = FALSE, alpha = 0.05)
  return(rel$results[2,])
}

bids_rel = get_ICC2(d_rel[,c("bids","bids_rel")]) # .79 [.55, .91]

pja_rel = get_ICC2(d_rel[,c("pja","pja_rel")]) # eps pass JA .34 [-.11, .67]

cja_rel = get_ICC2(d_rel[,c("cja","cja_rel")]) # eps coord JA .66 [.32, .85]

l_pja_rel = get_ICC2(d_rel[,c("l_pja","l_pja_rel")]) # pass JA length .33 [-.11, .67]

l_cja_rel = get_ICC2(d_rel[,c("l_cja","l_cja_rel")]) # coord JA length .60 [.24, .82]

save(d, file="Exp1_joint_attention_data.RData")
```

Inter-rater reliability is very high for number of bids, high for number of episodes of coordinated joint attention, but there is less agreement for episodes passive joint attention.

# Plots

Total parent bids (successful and unsuccessful) per minute.
```{r, echo=F, width=4, height=4}
ms_bids <- d %>%
  group_by(Condition) %>%
  multi_boot_standard(col = "bids_tot")

ms_bids$Condition <- factor(ms_bids$Condition,
levels = c("con", "exp"),
labels = c("control", "video"))

ggplot(ms_bids, aes(x = Condition, y = mean, fill = Condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Bids for Joint Attention per Min.") 
```

Total episodes of joint attention.

```{r, echo=F}
ms_tja <- d %>%
  group_by(Condition) %>%
  multi_boot_standard(col = "total_ja")

ms_tja$Condition <- factor(ms_tja$Condition,
levels = c("con", "exp"),
labels = c("control", "video"))

ggplot(ms_tja, aes(x = Condition, y = mean, fill = Condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Episodes of JA per Min.") 
```

Episodes of passive joint attention.
```{r, echo=F}
ms_pja <- d %>%
  group_by(Condition) %>%
  multi_boot_standard(col = "pja")

ggplot(ms_pja, aes(x = Condition, y = mean, fill = Condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") +  ylab("Episodes of Passive JA per Min.")  
```

Episodes of coordinated joint attention
```{r, echo=F}
ms_cja <- d %>%
  group_by(Condition) %>%
  multi_boot_standard(col = "cja")

ggplot(ms_cja, aes(x = Condition, y = mean, fill = Condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9)) +
  xlab("Condition") + ylab("Episodes of Coordinated JA per Min.")
```

Total duration of passive joint attention.
```{r, echo=F}
ms_lpja <- d %>%
  filter(!is.na(Condition))%>%
  group_by(Condition) %>%
  multi_boot_standard(col = "pja_length")

ggplot(ms_lpja, aes(x = Condition, y = mean, fill = Condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Seconds of Passive JA per Min.")
```

Total duration of coordinated joint attention.
```{r, echo=F}
ms_lcja <- d %>%
  filter(!is.na(Condition))%>%
  group_by(Condition) %>%
  multi_boot_standard(col = "cja_length")

ggplot(ms_lcja, aes(x = Condition, y = mean, fill = Condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Seconds of Coordinated JA per Min.")
```

## Demographics

### Child gender

Total bids by gender.
```{r, echo=F}
# total bids
ms_bids <- d %>%
  group_by(Condition, gender) %>%
  multi_boot_standard(col = "bids_tot")

ggplot(ms_bids, aes(x = Condition, y = mean, fill=gender)) + 
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Bids for Joint Attention per Min.") 
```

Total duration of passive joint attention by gender
```{r, echo=F}
ms_lpja <- d %>%
  filter(!is.na(Condition))%>%
  group_by(Condition, gender) %>%
  multi_boot_standard(col = "pja_length")

ggplot(ms_lpja, aes(x = Condition, y = mean, fill=gender)) + 
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Seconds of Passive JA per Min.")
```

Total duration of coordinated joint attention by gender
```{r, echo=F}
ms_lcja <- d %>%
  filter(!is.na(Condition))%>%
  group_by(Condition, gender) %>%
  multi_boot_standard(col = "cja_length")

ggplot(ms_lcja, aes(x = Condition, y = mean, fill=gender)) + 
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Seconds of Coordinated JA per Min.")
```

### Child age

Children were separated into 3 age groups: 6-12 months, 12-18 months, and 18-24 months.

Total bids by age.
```{r, echo=F}
# total bids
ms_bids <- d %>%
  group_by(Condition, age_group) %>%
  multi_boot_standard(col = "bids_tot")

ms_bids$age_group <- factor(ms_bids$age_group , levels = c("6","12", "18"))

ggplot(ms_bids, aes(x = Condition, y = mean, fill=age_group)) + 
  geom_bar(stat="identity", position= "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Bids for JA per Min.")
```

Total duration of passive joint attention by age
```{r, echo=F}
ms_lpja <- d %>%
  filter(!is.na(Condition))%>%
  group_by(Condition, age_group) %>%
  multi_boot_standard(col = "pja_length")

ms_lpja$age_group <- factor(ms_lpja$age_group , levels = c("6","12", "18"))

ggplot(ms_lpja, aes(x = Condition, y = mean, fill = age_group)) + 
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Seconds of Passive JA per Min.")
```

Total duration of coordinated joint attention by age
```{r, echo=F}
ms_lcja <- d %>%
  filter(!is.na(Condition))%>%
  group_by(Condition, age_group) %>%
  multi_boot_standard(col = "cja_length")

ms_lcja$age_group <- factor(ms_lcja$age_group , levels = c("6","12", "18"))

ggplot(ms_lcja, aes(x = Condition, y = mean, fill = age_group)) + 
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Seconds of Coordinated JA per Min.")
```

### Parenting attitudes and education

Do PAQ scores predict the number of bids for joint attention by parents?
```{r, echo=F}
d_plot <- d %>%
  gather("subscale","score", AA:RR)

ggplot(d_plot, aes(x = score, y = bids_tot, col = subscale)) +
  geom_jitter() + xlim(0,6) +
  geom_smooth(method="lm", se=FALSE)  + 
  facet_wrap(~Condition)
```

Does parent education predict the number of bids for joint attention by parents?
```{r, echo=F}
d_plot <- d %>%
  gather("subscale","score", AA:RR)

ggplot(d_plot, aes(x = parent_ed, y = bids_tot)) +
  geom_jitter() + 
  xlim(0,6) +
  geom_smooth(method="lm", se=FALSE)  + 
  facet_wrap(~Condition)
```

# Analysis

Is the duration of joint attention correlated with age in months?

```{r}
cormat <- d %>%
  select(age_months, pja_length, cja_length, total_lja)
rcorr(as.matrix(cormat))
```

## Mixed effects models.

Prepare data.
```{r}
lmer_data <- d %>%
  filter(!is.na(AA), !is.na(EL), !is.na(RR))%>%
  mutate(Condition = factor(Condition), 
         bids_tot = as.numeric(bids_tot),
         EL = as.numeric(langcog::scale(EL, scale=FALSE)),
         AA = as.numeric(langcog::scale(AA, scale=FALSE)),
         RR = as.numeric(langcog::scale(RR, scale=FALSE)),
         age = as.numeric(langcog::scale(age, scale=FALSE)),
         gender = as.factor(gender),
         Video = as.factor(Video))
```

Total number of bids
```{r}
maximal_mod <- lmer(bids_tot ~ Condition *  EL  + Condition * AA + Condition * RR + age + gender + parent_ed +
                           (1| Video), data = lmer_data)
summary(maximal_mod)
```

Episodes of coordinated joint attention.
```{r}
maximal_mod <- lmer(cja ~ Condition *  EL  + Condition * AA + Condition * RR  + age + gender + parent_ed +
                           (1| Video), data = lmer_data)
summary(maximal_mod)
```

Episodes of passive joint attention.
```{r}
maximal_mod <- lmer(pja ~ Condition *  EL  + Condition * AA + Condition * RR  + age + gender + parent_ed +
                           (1| Video), data = lmer_data)
summary(maximal_mod)
```

Total duration of passive joint attention.
```{r}
maximal_mod <- lmer(pja_length ~ Condition *  EL  + Condition * AA + Condition * RR  + age + gender + parent_ed +
                           (1| Video), data = lmer_data)
summary(maximal_mod)
```

Total duration of coordinated joint attention.
```{r}
maximal_mod <- lmer(cja_length ~ Condition *  EL  + Condition * AA + Condition * RR  + age + gender + parent_ed +
                           (1| Video), data = lmer_data)
summary(maximal_mod)
```

# Conclusions

There is a main effect of condition on total bids for joint attention. Parents in the experimental condition (i.e., those who saw a video demonstrating an activity) made a greater number of bids for joint attention with their child.

There was no effect of condition on the number of episodes of either passive or coordinated joint attention, or the duration of these episodes.

There is a marginal effect of gender on bids for joint attention, with parents of males producing more bids. There is a marginal interaction between RR scores and condition on passive joint attention, such that the experimental condition increased the number of episodes of PJA to a greater extent for people with high RR scores.

While the electronically-delivered parenting advice increased the number of bids for joint attention by parents, it did not significantly effect the number or duration of episodes of joint attention. One possibility is that child variables had a comparatively larger impact on the attainment of joint attention.