### 2020-03-16 parenting_obs_e2/joint_attention/parenting_obs_JA.Rmd

---
title: "Parenting Observational Study E2 Coding"
author: "Emily, Vivian, George"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    number_sections: true
---

<style type="text/css">
body, td {
   font-size: 14px;
}
code {
  font-size: 11px;
}
pre {
  font-size: 11px;
}
</style>

Looking at the effect of electronically-delivered parenting interventions on parent-child interactions. Parents in the experimental condition viewed a video (designed for a parenting app) demonstrating an activity they could do with their child. Parents in the control condition did not view a video. Both groups of parents were asked to play with their child for 3 minutes with a set of props that matched the video demonstration they had seen. Children were 6-24 months of age. There were 6 videos total (2 per age group). Assignment to videos/sets of toys was counterbalanced within age groups. Parents also completed the Parenting Attitudes Questionnaire. Videos of the play sessions were coded for bids for joint attention, episodes of passive joint attention, and episodes of coordinated joint attention.

# Data preprocessing

Preliminaries.

```{r include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(fig.width=8, fig.height=5, 
                      echo=TRUE, warning=FALSE, message=FALSE, cache=TRUE)
library(tidyverse)
library(lme4)
library(knitr)
library(markdown)
library(lmerTest)
library(stringr)
library(langcog)
library(xtable)
library(Hmisc)
theme_set(ggthemes::theme_few() + theme(legend.title = element_text(size=18), legend.text = element_text(size=16), 
        axis.text.x  = element_text(vjust=0.5, size=16), axis.text.y  = element_text(vjust=0.5, size=16),
        axis.title.x = element_text(size=18), axis.title.y = element_text(size=18))) 
```

Read in files and consolidate to the same directory.

```{r echo=F}
# for rate coding (JA/minute), we coded the start / end of play sessions because they were not exactly 3 minutes
play_duration <- read.csv("parenting_obs_play_duration_E2.csv", header=T)
play_duration <- play_duration %>% 
  mutate(play_duration = play_stop_sec - play_start_sec)
sort(play_duration$play_duration) # 1 pretty short session: 131 seconds
summary(play_duration$play_duration) # median = 184.5, mean=185.8
# use play_duration$verbalcue_start_sec instead of play_start? (when it exists)

combine_datavyu_files <- function(data_dir) {
  d.long <- data.frame()
  files <- dir(data_dir)

  for (f in files) {
    jf <- paste(data_dir, f, sep="")
    jd <- read.csv(jf)

    # (assumes something happened early in the vid and late in the vid..)
    #vid_duration = jd[nrow(jd),]$JointAttention.offset - jd[1,]$JointAttention.onset
    #if(vid_duration < 180000) print(paste(f, "final - first coded time is less than 180 s:",vid_duration))
    SID = str_replace(f, ".csv", "")
    vid_duration = play_duration[which(play_duration$sid==SID),]$play_duration

    jd$phase_length = jd$JointAttention.offset - jd$JointAttention.onset
    if(length(which(jd$phase_length < 0))!=0) print(paste(f, "episode with negative duration!"))

    id <- data.frame(SID = SID, 
                   line = jd$JointAttention.ordinal, 
                   phase_length = jd$phase_length,
                   pja = jd$JointAttention.pJA, # was episode passive?
                   cja = jd$JointAttention.cJA,
                   bid = jd$JointAttention.bids, # parent bids (1=parent, 0=kid)
                   vid_duration = vid_duration
                )

    d.long <- rbind(d.long, id)
    #d.raw <- bind_rows(d.raw, sd)
  }
  return(d.long)
}

d.long <- combine_datavyu_files("data/")

d.long[which(d.long$phase_length<0),] # SID 121317_1  line 17          -68   1   0   1          181

d.long[which(d.long$phase_length<3000),c("cja","pja")] = 0 # any episodes <3s not counted (except bids)
# 337 / 1094

# summarise per subject
d.raw <- d.long %>% group_by(SID) %>%
  mutate(vid_duration = vid_duration / 60, # seconds to minutes
         phase_length = phase_length / 1000) %>% # ms to s
  summarise(bids=sum(bid), pja_length=sum(pja*phase_length), cja_length=sum(cja*phase_length),
            pja=sum(pja), cja=sum(cja), vid_duration=vid_duration[1]) %>%
  mutate(total_ja = pja + cja, total_lja = pja_length + cja_length)

# normalize to rates (e.g., bids/minute and seconds / min duration of JA)
vid_dur = d.raw$vid_duration
d.raw[,2:9] = d.raw[,2:9] / d.raw$vid_duration
d.raw$vid_duration = vid_dur
```

Read in trial info 
```{r}
conditions <- read.csv("conditions.csv") %>%
  transmute(SID = sid, condition=condition, video=video, label=label)

d <- left_join(d.raw, conditions, by="SID")

#d$age[d$sid=="111517_3"] = "2"
```

# Reliability

Import reliability coder's data (25 participants were coded for reliability).
```{r}
d.long_rel <- combine_datavyu_files("rel_data/")

d.long_rel[which(d.long_rel$phase_length<3000),c("cja","pja")] = 0 # any episodes <3s not counted (except bids)

# summarise per subject
d.raw_rel <- d.long_rel %>% group_by(SID) %>%
  mutate(vid_duration = vid_duration / 60, # seconds to minutes
         phase_length = phase_length / 1000) %>% # ms to s
  summarise(bids=sum(bid), pja_length=sum(pja*phase_length), cja_length=sum(cja*phase_length),
            pja=sum(pja), cja=sum(cja), vid_duration=vid_duration[1]) %>%
  mutate(total_ja = pja + cja, total_lja = pja_length + cja_length)

# normalize to rates (e.g., bids/minute)
d.raw_rel[,2:9] = d.raw_rel[,2:9] / d.raw_rel$vid_duration
```

Reliability.
```{r}
d.raw_rel <- d.raw_rel %>%
  mutate(bids_rel = bids, pja_rel = pja, cja_rel = cja, l_pja_rel = pja_length, l_cja_rel = cja_length) %>%
  select(SID, bids_rel, pja_rel, cja_rel, l_cja_rel, l_pja_rel)

d.raw <- d.raw %>% mutate(l_pja = pja_length, l_cja = cja_length)
# 110416_03 l_cja_rel = -25889, which shouldn't be possible (incomplete final row of file)
# 121516_01 has NA bids_rel, pja_rel, and cja_rel..
# 111716_08 l_cja = -49923, which shouldn't be possible

# 110416_05, 050116_01, and 092216_03: reliability coder saw no passive JA
d_rel <- d.raw_rel %>% left_join(d.raw, by="SID") %>%
  select(SID, bids, bids_rel, pja, pja_rel, cja, cja_rel, l_cja, l_cja_rel, l_pja, l_pja_rel)

get_ICC2 <- function(dat)  {
  rel = psych::ICC(dat, missing = FALSE, alpha = 0.05)
  return(rel$results[2,])
}

bids_rel = get_ICC2(d_rel[,c("bids","bids_rel")]) # .81 [.62,.91]

pja_rel = get_ICC2(d_rel[,c("pja","pja_rel")]) # .74 eps pass JA [.48,.88]

cja_rel = get_ICC2(d_rel[,c("cja","cja_rel")]) # .80 eps coord JA [.61,.91]

l_pja_rel = get_ICC2(d_rel[,c("l_pja","l_pja_rel")]) # .72 pass JA length [.44,.86]

l_cja_rel = get_ICC2(d_rel[,c("l_cja","l_cja_rel")]) # .88 coord JA length [.75,.94]
```

Inter-rater reliability is very high for length of coordinated JA (`r round(l_cja_rel$ICC, 2)`), high for bids (`r round(bids_rel$ICC, 2)`) and episodes of coordinated JA (`r round(cja_rel$ICC, 2)`), and not bad for episodes (`r round(pja_rel$ICC, 2)`) and duration of passive JA (`r round(l_pja_rel$ICC, 2)`).

# Plots

Total number of bids (successful and unsuccessful)
```{r, width=4, height=4}
dem <- read.csv("../anonymized_demographics_parenting_obs_e2.csv", header=T)
dem$age = as.numeric(as.character(dem$age))
d$age = NULL
d <- d %>% left_join(dem, by="SID") %>%
  filter(age <= 2.0) # 1 kid exactly 2...

save(d, file="Exp2_joint_attention_data.RData")

d$condition <- factor(d$condition, levels = c("con", "exp"), labels = c("control", "video"))

ms_bids <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "bids")

ggplot(ms_bids, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Bids for Joint Attention per Min.") +
  langcog::scale_colour_solarized() + theme(legend.position="none") 
```

Total episodes of joint attention.

```{r, echo=F}
ms_tja <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "total_ja")

ggplot(ms_tja, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Episodes of JA per Min.") 
```

Episodes of passive joint attention.
```{r, echo=F}
ms_pja <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "pja")

ggplot(ms_pja, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Episodes of Passive JA per Min.") 
```

Episodes of coordinated joint attention
```{r, echo=F}
ms_cja <- d %>%
  group_by(condition) %>%
  multi_boot_standard(col = "cja")

ggplot(ms_cja, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + 
  ylab("Episodes of Coordinated JA per Min.") 
```

Rate of passive joint attention.
```{r, echo=F}
ms_lpja <- d %>%
  filter(!is.na(condition))%>%
  group_by(condition) %>%
  multi_boot_standard(col = "pja_length")

ggplot(ms_lpja, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Seconds of Passive JA per Min.") 
```

Rate of coordinated joint attention.
```{r, echo=F}
ms_lcja <- d %>%
  filter(!is.na(condition))%>%
  group_by(condition) %>%
  multi_boot_standard(col = "cja_length")

ggplot(ms_lcja, aes(x = condition, y = mean, fill = condition)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Seconds of Coordinated JA per Min.")
```

## Demographics

### Child gender

Total bids by gender.

```{r, echo=F}
# total bids
ms_bids <- d %>%
  group_by(condition, gender) %>%
  multi_boot_standard(col = "bids")

ggplot(ms_bids, aes(x = condition, y = mean, fill=gender)) + 
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9)) +
  xlab("Condition") + ylab("Bids for Joint Attention per Min.") # look up missing gender: "102518_6" "110418_3"
```

Rate of passive joint attention by gender
```{r, echo=F}
ms_lpja <- d %>%
  filter(!is.na(condition))%>%
  group_by(condition, gender) %>%
  multi_boot_standard(col = "pja_length")

ggplot(ms_lpja, aes(x = condition, y = mean, fill=gender)) + 
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Seconds of Passive JA per Min.") 
```

Rate of coordinated joint attention by gender

```{r, echo=F}
ms_lcja <- d %>%
  filter(!is.na(condition))%>%
  group_by(condition, gender) %>%
  multi_boot_standard(col = "cja_length")

ggplot(ms_lcja, aes(x = condition, y = mean, fill=gender)) + 
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Seconds of Coordinated JA per Min.") 
```

### Child age

Children were separated into 3 age groups: 6-12 months, 12-18 months, and 18-24 months.

Total bids by age.

```{r, echo=F}
# total bids
d$age_group = ifelse(d$age<1, "6-12", ifelse(d$age<1.5, "12-18", "18-24"))
d$age_group <- factor(d$age_group , levels = c("6-12","12-18", "18-24"))

ms_bids <- d %>%
  group_by(condition, age_group) %>%
  multi_boot_standard(col = "bids")

ggplot(ms_bids, aes(x = condition, y = mean, fill=age_group)) + 
  geom_bar(stat="identity", position= "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Bids for Joint Attention per Min.") 
```

Rate of passive joint attention by age

```{r, echo=F}
ms_lpja <- d %>%
  filter(!is.na(condition))%>%
  group_by(condition, age_group) %>%
  multi_boot_standard(col = "pja_length")

ggplot(ms_lpja, aes(x = condition, y = mean, fill = age_group)) + 
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Seconds of Passive JA per Min.") 
```

Rate of coordinated joint attention by age

```{r, echo=F}
ms_lcja <- d %>%
  filter(!is.na(condition))%>%
  group_by(condition, age_group) %>%
  multi_boot_standard(col = "cja_length")

ggplot(ms_lcja, aes(x = condition, y = mean, fill = age_group)) + 
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+
  xlab("Condition") + ylab("Seconds of Coordinated JA per Min.") 
```

# Analysis

Is the duration of joint attention correlated with age?

```{r, echo=F}
cormat <- d %>%
  select(age, pja_length, cja_length, total_lja)
rcorr(as.matrix(cormat))
```

Mixed effects models.

Prepare data.
```{r}
lmer_data <- d %>% 
  mutate(Condition = factor(condition), 
         bids = as.numeric(bids),
         age = as.numeric(langcog::scale(age, scale=FALSE)),
         gender = as.factor(gender),
         Video = as.factor(video))
```

Total number of bids
```{r}
lmer_data$parent_ed = as.numeric(as.character(lmer_data$parent_ed))
maximal_mod <- lmer(bids ~ Condition * age + (1| Video), data = lmer_data)
summary(maximal_mod)
```

Episodes of coordinated joint attention.
```{r}
maximal_mod <- lmer(cja ~ Condition * age + (1| Video), data = lmer_data)
summary(maximal_mod)
```

Episodes of passive joint attention.
```{r}
maximal_mod <- lmer(pja ~ Condition * age + (1| Video), data = lmer_data)
summary(maximal_mod)
```

Total duration of passive joint attention.
```{r}
maximal_mod <- lmer(pja_length ~ Condition * age + (1| Video), data = lmer_data)
summary(maximal_mod)
```

Total duration of coordinated joint attention.
```{r}
maximal_mod <- lmer(cja_length ~ Condition * age + (1| Video), data = lmer_data)
summary(maximal_mod)
```

# Conclusions

As in Exp. 1, there is a significant effect of video condition on parental bids for JA: parents made more bids after watching an activity video. There was also a significant effect of video condition on the number of episodes of coordinated JA, as well as an interaction of condition and age.

There was a significant effect of age on the rate of episodes and duration of passive JA, and on the duration of coordinated JA.