### 2020-04-08 paper/appendix-Bayesian-analyses.Rmd
fixed a couple NAs and line 328 issue, and improved appendix

---
title: 'Appendix: Bayesian Analysis Overview'
output: pdf_document
---

```{r setup, include = FALSE}
suppressPackageStartupMessages(c("dplyr","langcog","tidyr","ggplot2","lme4"))
library("papaja")
library(png)
library(grid)
library(ggplot2)
library(xtable)
library(psych)
library(langcog)
library(tidyverse)
library(ggthemes)
library(lme4)
library(lmerTest)
library(tidyboot)
library(knitr)
library(brms)
library(rstanarm)
library(bayestestR)
library(bayesplot)

source("multiplot.R")
select <- dplyr::select # masked by MASS
theme_set(theme_few())

# labels=addline_format(c("Activity Video", "Science Video"))
addline_format <- function(x,...){
    gsub('\\s','\n',x)
}

get_stan_glmer_reporting_values <- function(model, cond, digits=3) {
  ci95 <- posterior_interval(model, prob = 0.95, pars = cond)
  ci89 <- posterior_interval(model, prob = 0.89, pars = cond) # 89% CIs as recommended by Kruschke and others
  pd = p_direction(model)
  pd = pd[which(pd$Parameter==cond),]$pd
  ci = round(c(ci95,ci89,pd), digits)
  names(ci) = c("95pLB","95pUB","89pLB","89pUB","pd") # 95% and 89% lower and upper bounds
  beta = round(model$coefficients[cond], digits)
  return(c(beta,ci))
}

# pars are specific to E1
plot_posterior <- function(mod, title="") {
  posterior <- as.matrix(mod)
  plot_title <- ggtitle(title)
  mcmc_areas(posterior,
           pars = c("conditionexp", 
                    "EL", "AA", "RR", 
                    "age", "genderM", "parent_ed", 
                    "conditionexp:EL",
                    "conditionexp:AA",
                    "conditionexp:RR"),
           prob = 0.89, prob_outer = 0.95, point_est = "mean") + 
           theme(plot.margin=unit(c(-25,2,2,2),"mm")) +
           annotate("text", x = -0, y = 10.63, label = title)
}
```

```{r analysis-preferences, echo=F}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r load-exp1-lexical-data, include=F}
e1path = "../parenting_obs_e1/"

d = read.csv("../language_analyses/exp1_lexical_diversity.csv")

e1_parent_ed = table(d$parent_ed)

e1ld <- d %>%
  mutate(condition = factor(condition), 
         gender = as.factor(gender),
         video = as.factor(video))

e1_gender = table(e1ld$gender)

# rate calculations (tokens / min, words / min)
e1_play_duration <- read.csv(paste0(e1path,"joint_attention/parenting_obs_play_duration_E1.csv"), header=T)
e1_play_duration$start_sec = with(e1_play_duration, ifelse(is.na(verbalcue_start_sec), 
                                                           play_start_sec, verbalcue_start_sec))
e1_play_duration <- e1_play_duration %>% 
  mutate(play_duration = (play_stop_sec - start_sec) / 60) %>% # sec to mins
  select(SID, play_duration)
e1ld <- e1ld %>% left_join(e1_play_duration, by=c("sid"="SID"))

# convert types and tokens to rate: types/min and tokens/min
e1ld <- e1ld %>% 
  mutate(types = (types / play_duration), 
         tokens = (tokens / play_duration),
         TTR = (TTR / play_duration),
         MTLD = (MTLD / play_duration))

e1model_string = "~ condition*EL + condition*AA + condition*RR  + age + gender + parent_ed + (1|video)"
```

Here we look at one of the regressions from Experiment 1 in detail in order to explain and illustrate the rationale of the Bayesian analysis, in particular the *probability of direction* (*pd*) and the 89% CI (credible interval). For this example we take the exploratory regression from Experiment 1 predicting parents' rate of word types with fixed effects of condition (*No Video* or *Activity Video*), each subscale of the EPAQ (EL: Early Learning, AA: Affection and Attachment, and RR: Rules & Respect), parent education (parent_ed), the child's age (centered) and gender. The model also included interactions for condition and each subscale of the EPAQ, as well as a random intecept for each video.
The model syntax was: types \~ condition\*EL + condition\*AA + condition\*RR  + age + gender + parent_ed + (1|video).
Table 1 reports the mean of each coefficient's posterior distribution, along with the lower and upper bounds for the 89% credible interval, and the probability of direction (*pd*).

```{r, run-regression, echo=F, include=F}
types_modE <- stan_glmer(paste("types", e1model_string), data=e1ld) 
summary(types_modE) 
diagnostic_posterior(types_modE)

e1ed = get_stan_glmer_reporting_values(types_modE, 'parent_ed') 
pd = p_direction(types_modE, effects="fixed") # cond 82.5% EL 82.5% AA 87% genderM 93.2%, parent_ed 93.5% cond:AA 94.5%

tok_modE1 <- stan_glmer(paste("tokens", e1model_string), data=e1ld) 
mtld_modE1 <- stan_glmer(paste("MTLD", e1model_string), data=e1ld) 
ttr_modE1 <- stan_glmer(paste("TTR", e1model_string), data=e1ld) 
```

```{r, p-direction, echo=F, fig.cap="Mean, *pd*, and 89% CIs of the coefficient posterior distributions."}

cols = c("Mean","89% CI Lower","89% CI Upper")
pd[,cols] = NA
for(i in 2:nrow(pd)) {
  tmp = get_stan_glmer_reporting_values(types_modE, pd[i,]$Parameter) 
  pd[i,cols] = c(tmp[pd[i,]$Parameter], tmp["89pLB"], tmp["89pUB"])
}
kable(pd[2:nrow(pd),c("Parameter","pd", cols)], digits=2, row.names=F)
```

What do these values mean? 
Let's take a closer look at the posterior distribution of the parent education (parent_ed) coefficient. 
Shown below in Figure 1, the 87% of this distribution that is greater than 0 is shaded orange, which corresponds to the probability of direction *pd* = 0.87, which has the straightforward interpretation of being the probability that an effect has the same sign as the median value of the posterior. Thus, *pd* has a range of 0.5 (if the distribution is equally-distributed around 0: a likely true null effect) to 1.0 (if the posterior is entirely positive or negative).
Since *pd* < 0.95 in this case (corresponding to a $p$<.05 threshold in the null hypothesis significance-testing framework), we do not consider parent education to have a notable effect on the rate of types used during play, but the posterior distribution is interpretable.
Also shown are the mean estimated coefficient value (red line) and 89% credible intervals (blue lines), within which 89% of the posterior distribution falls.

```{r, parent-ed-effect, echo=F, fig.cap="Posterior distribution of parent education coefficient with mean (red), 89% CIs (blue), and *pd* (orange portion) shown.", fig.width=4, fig.height=2.0}
dat = types_modE %>% posterior_samples("parent_ed") %>%
  estimate_density(extend=TRUE)

dat %>% 
  ggplot(aes(x=x, y=y)) +
  geom_area(fill="grey") +
  geom_ribbon(data=subset(dat, x>0),aes(ymax=y),ymin=0,
              fill="orange",colour=NA,alpha=0.6) + 
  theme_classic() + ylab("Posterior Density") + xlab("Candidate Values of Parent Education Coefficient") +
  geom_vline(xintercept=e1ed["89pLB"], color="royalblue", size=1) +
  geom_vline(xintercept=e1ed["89pUB"], color="royalblue", size=1) +
  geom_vline(xintercept=0, size=1, lty="dashed") + 
  geom_vline(xintercept=e1ed["parent_ed"], size=1, color="red")
```

Figure 2 shows the posteriors for all of the coefficients in this regression. 
Only the interaction between condition and the Affection and Attachment subscale of the EPAQ was notable (by *pd* in Table 1), but as indicated by their *pd* values, much of the mass some other parameters' posteriors are skewed negative (e.g., conditionexp: the Activity Video condition; EL: the Early Learning subscale of the EPAQ), while others are more evenly-distributed (e.g., RR, $pd$ = 0.51).

```{r, plot-posteriors, echo=F, fig.width=6, fig.height=6, fig.cap="Posterior distributions with means and 89% credible intervals. The outermost 5% of each posterior distribution is clipped and not shown."}
plot_posterior(types_modE, title="Exp. 1: Word Types")
```

```{r, plot-token-posterior, echo=F, fig.width=6, fig.height=6, fig.cap="Posterior distributions with means and 89% credible intervals. The outermost 5% of each posterior distribution is clipped and not shown."}

#e1ed = get_stan_glmer_reporting_values(tok_modE1, 'parent_ed') 
#pd = p_direction(tok_modE1, effects="fixed") # cond 82.5% EL 82.5% AA 87% genderM 93.2%, parent_ed 93.5% cond:AA 94.5%

plot_posterior(tok_modE1, "Exp. 1: Word Tokens")

```

```{r, echo=F, fig.width=6, fig.height=6, fig.cap="Posterior distributions for regression predicting TTR, with means and 89% credible intervals. The outermost 5% of each posterior distribution is clipped and not shown.", warning=F}
plot_posterior(ttr_modE1, "Exp. 1: TTR")
```

```{r, echo=F, fig.width=6, fig.height=6, fig.cap="Posterior distributions for regression predicting MTLD, with means and 89% credible intervals. The outermost 5% of each posterior distribution is clipped and not shown."}
plot_posterior(mtld_modE1, "Exp. 1: MTLD")
```